---
title: Minimal Matrix Problem
format: html
---

# Setup


```{r}
library(tidyverse)
```


```{r}
combs <- readRDS("./data/combs.rds")
mat <- readRDS("./data/mat.rds")
correct_res <- readRDS("./data/res.rds")
mat_1 <- mat - 1L
```

```{r}
options("matprod")
```

> a string selecting the implementation of the matrix products %*%, crossprod, and tcrossprod for double and complex vectors: 

> "default":
> uses BLAS to speed up computation, but to ensure correct propagation of NaN and Inf values it uses an unoptimized 3-loop algorithm for inputs that may contain NaN or Inf values. When deemed beneficial for performance, "default" may call the 3-loop algorithm unconditionally, i.e., without checking the input for NaN/Inf values. The 3-loop algorithm uses (only) a double accumulator for summation, which is consistent with the reference BLAS implementation.


```{r}
class(mat_1[1])
class(combs[1])
```

```{r}
N  <- 10
outputs = vector("list", N)
for (i in 1:N) {
  res <- mat_1 %*% combs
  outputs[[i]] <- res
}

map2_lgl(outputs, lag(outputs), identical)
```

```{r}
options(matprod = "internal")
```

> "internal"
> uses an unoptimized 3-loop algorithm which correctly propagates NaN and Inf values and is consistent in precision with other summation algorithms inside R like sum or colSums (which now means that it uses a long double accumulator for summation if available and enabled, see capabilities).

```{r}
N  <- 10
outputs = vector("list", N)
for (i in 1:N) {
  res <- mat_1 %*% combs
  outputs[[i]] <- res
}

map2_lgl(outputs, lag(outputs), identical)
```


```{r}
capabilities()
```


```{r}
options(matprod = "blas")
```


```{r}
N  <- 10
outputs = vector("list", N)
for (i in 1:N) {
  res <- mat_1 %*% combs
  outputs[[i]] <- res
}

map2_lgl(outputs, lag(outputs), identical)
```


```{r}
any(is.nan(mat))
```

```{r}
any(is.na(mat))
```

```{r}
any(is.infinite(mat))
```

```{r}
any(is.nan(combs))
```

```{r}
any(is.na(combs))
```

```{r}
any(is.infinite(combs))
```


```{r}
options(matprod = "default")
```


